Weighted Visibly Pushdown Automata and Automated Music Transcription

### 2021-04-27 main.tex
def. transducers and automata

\begin{definition}
A weighted transducer $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
and the semiring $\Semiring$ is a tuple
$T = \< Q, \init, \weight, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
is a function defining the weight for entering, 
respectively leaving, a state, 
and $\weight$ is a transition %weight 
function of $Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
\end{definition}
We extend the above transition function into a function from
$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
into $\Semiring$, also called $\weight$ for simplicity, such that 
for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
and with 
$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \weight(q, q')$, 
\[
\begin{array}{rcl}
\weight(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
\weight(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
\weight(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
\weight(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
\end{array}      
\]
%More precisely, $Q \times Q$, 
%resp. $Q \times \Sigma \times Q$,
%$Q \times \Delta \times Q$,
%$Q \times \Sigma \times \Delta \times Q$,
%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.

@@

>>>>

\begin{definition}
A \emph{weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
and the semiring $\Semiring$ is a tuple
$T = \< Q, \init, \weight, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
is a function defining the weight for entering, 
respectively leaving, a state, 
and $\weight$ is a transition %weight 
function of $Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
\end{definition}
We extend the above transition function into a function from
$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
into $\Semiring$, also called $\weight$ for simplicity, such that 
for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
and with 
$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \weight(q, q')$, 
\[
\begin{array}{rcl}
\weight(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
\weight(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
\weight(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
\weight(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
\end{array}      
\]
%More precisely, $Q \times Q$, 
%resp. $Q \times \Sigma \times Q$,
%$Q \times \Delta \times Q$,
%$Q \times \Sigma \times \Delta \times Q$,
%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.

@@

The weighted transducer $T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into the weights of~$\Semiring$,
based on the following intermediate function $\weight_\A$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_\A(q, s, t, q') & = & & \weight(q, \epsilon, \epsilon, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                      \scriptstyle q'' \in Q\\[-2pt]
                                      \scriptstyle s = au, a \in \Sigma
                                      \end{array}} &
     \weight(q, a, \epsilon, q'') \otimes \weight_\A(q'', u, t, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
      \weight(q, \epsilon, b, q'') \otimes \weight_\A(q'', s, v, q')\\ 
      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                         \scriptstyle q'' \in Q\\[-2pt]
                                         \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                         \scriptstyle t = bv, b \in \Delta
                                         \end{array}} &
      \weight(q, a, b, q'') \otimes \weight_\A(q'', u, v, q')\\ 
\end{array}
\]
Recall that by convention, an empty sum with $\oplus$ is $\zero$.

\noindent
The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
is then defined as follows: 
\[
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_\A(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
\]

A \emph{weighted automata} $T = \< Q, \init, \weight, \final >$
over $\Sigma$ and $\Semiring$ 
is defined in a similar way by simply omitting the output symbols,
\ie $\weight$ is a function of~$Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma >$, 
or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
